A new simplified manual tour, with examples in mathematica and R

### 2022-04-12 paper/paper.Rmd
In-plane rotation idea seems to be working

>>>>

```{r }
source("../code/plot_basis.R")
source("../code/linear_alg.R")

# Base plot set up
plot.new()
par(pty="s", xaxt="n", yaxt="n", bty="n",
      omi=c(0,0,0,0), mar=c(0,0,0,0),
    mfrow=c(1,4))

# j is the component that should be kept fixed
orthonormalize_rotated <- function(A, j){
  m1 <- A[j, 1]
  m2 <- A[j, 2]
  c <- 1 / (sqrt(1 + (m2/m1)^2))
  s <- (m2/m1) / (sqrt(1 + (m2/m1)^2))
  rot <- matrix(c(c, -s, s, c), nrow = 2)
  rot_inv <- matrix(c(c, s, -s, c), nrow = 2)
  for(i in 1:nrow(A)){
    A[i,] <- rot %*% A[i,]
  }
  A[,1] <- normalize_w_fixed(A[,1], j)
  A <- tourr::orthonormalise(A)
  for(i in 1:nrow(A)){
    A[i,] <- rot_inv %*% A[i,]
    }
  A
}

# normalize a vector while keeping component j fixed
normalize_w_fixed <- function(v, j){
  c <- v[j]
  k <- sqrt((1-c^2) / sum(v[-j]^2))
  v[-j] <- k * v[-j]
  v
}

# Iterate and save images
# Unconstrained orthonormalisation
Anew <- A
for (i in 1:3) {
  Anew[vchange,] <- A[vchange,]*(1-i*eps)
  exact <- Anew[vchange,]
  Anew <- orthonormalize_rotated(Anew, vchange)
  plot_basis(Anew, vchange, vcol="black")
  points(exact[1], exact[2], pch=16, col="orange")
}

```

### 2022-04-12 paper/paper.tex
title change
In-plane rotation idea seems to be working

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

\title{A new simplified manual tour, with examples in mathematica}

@@

>>>>

% see https://stackoverflow.com/a/47122900

% Pandoc citation processing

@@

\title{A new simplified manual tour, with examples in mathematica and R}

@@

\hypertarget{potential-algorithm-3}{%
\subsubsection{Potential Algorithm 3}\label{potential-algorithm-3}}

For now just sketching the idea:

\begin{itemize}
\tightlist
\item
  first step is to click on the axis display to change the contribution
  of one variable \(m\)
\item
  capture that position and replace the corresponding row in the
  projection matrix
\item
  for 2D projection this gives components \(m_1\) along ``x'' direction
  and \(m_2\) along \(y\) direction
\item
  next step: rotate basis such that direction of row m now corresponds
  to the first basis vector, this means we apply 2x2 rotation matrix to
  each row of the projection matrix, where the rotation angle is
  \(tan(\theta) = m_2/m_1\) (that matrix can be written just in terms of
  that ratio, \(x= m_2/m_1\), when \(m_1<0\) I think we just need to
  translate \(\theta \rightarrow \theta + \pi\))
\item
  take the rotated basis, apply Gram-Schmidt, now the direction of \(m\)
  will not change (but the length will change during normalization)
\item
  rotated back to the original xy basis to update the plots
\end{itemize}

\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
